SoK: Defenses Against Web Application Attacks: From Research Approaches to Practical Tools

### 2014-11-13 docs/questioning.tex
changed title
pass in intro

Web application attacks may involve security misconfigurations, broken
authentication and session management, or other issues. Some of the
most dangerous web application attacks are associated with the
processing of external input data that can contain malicious code, a
vulnerability known as code injection.

Such attacks include Cross-Site Scripting ({\sc xss})
attacks~\cite{SG07}, {\sc sql} injection attacks~\cite{RL12b},
Cross-Site Request Forgery ({\sc csrf}) attacks~\cite{LZRL09}, and
others. Attacks like these have been topping the vulnerability lists
of numerous bulletin providers like the Open Web Application Security
Project\footnote{\url{https://www.owasp.org/index.php/Category:OWASP_Top_Ten_Project}}
({\sc owasp}) and {\sc
  mitre}\footnote{\url{http://cwe.mitre.org/top25/}} for several
years. Consider the case of {\sc owasp}'s Top Ten project whose main
goal is to raise awareness about web application security by
identifying some of the most critical risks facing organizations. The
Top Ten project is popular, referenced by numerous researchers. In its
three consecutive Top Ten lists (2007, 2010, 2013), different source
code-driven injection attacks dominate the top five positions.
Malicious users find new ways~\cite{HNSHS12,DKH14} to bypass defense
mechanisms by using a variety of techniques despite the numerous
countermeasures that are being introduced. The number of systems
developed to counter {\sc sql} injection attacks until 2006 was more
than twenty~\cite{HVO06}. Since then the number has doubled.
Researchers have indicated that the number of such attacks is steadily
increasing in recent years~\cite{SSL12}.

In this paper we show how the above attacks can be seen from a {\it
  common perspective}. We achieve this by providing an exploitation
model that highlights that the majority of the steps needed to perform
application attacks based on source code injection are
common.\footnote{Note that we leave out attacks that are based on
  binary code injection (i.e. buffer overflows --- see
  Section~\ref{sec:attacks}).} This is validated by the fact that
there are protection mechanisms that can detect more than one of these
attacks. Then, we {\it categorize} a selection of these protection
mechanisms. In the selection we included protection mechanisms that
detect such attacks and are presented in publications cited more than
twenty times. We also added some recent mechanisms that have been
presented in top security conferences, even though they have been
cited less than twenty times. We included the second group so that we
would not penalize recent research. Finally, we {\it analyze} each
mechanism based on the following dimensions:
\begin{itemize}
\item {\it Accuracy} Protection mechanisms are as good
  as their detection capability; this requires low false positive and
  false negative rates.
\item {\it Availability} We check if the mechanism and its
  testbed are available on-line.
\item {\it Computational Performance} We examine the computational
  overhead of the mechanisms at their points of deployment.
\item {\it Ease of use} We examine if the protection
  mechanism is practical in terms of deployment
  and can be easily adopted by security experts.
\item {\it Security} We examine how resilient is the system to
  attackers that know how it works and wants to circumvent it.
\item {\it Point of Detection} Based on our exploitation model
  we examine the location where a mechanism detects an attack.
\end{itemize}

\noindent
All the aforementioned requirements are considered critical
when building security mechanisms that protect
applications~\cite{A01,A00,SPWS13,nature2014}.
Based on this analysis we can identify the advantages
and disadvantages of the
various mechanisms and enumerate some of their common
characteristics. We can also draw useful conclusions
about the various categories and see how they compare
to each other.

There is already a survey on mitigating software vulnerabilities
in general~\cite{SZ12}. The scope of that
research is pretty wide and leaves out the majority of the
approaches and mechanisms that we report here.
This is mainly because that work includes countermeasures
that detect such vulnerabilities via program analysis
(which is done during development or testing), while
we include protection mechanisms that counter
such attacks when they take place. Static or
dynamic analyzers that examine applications to
identify vulnerabilities that may lead to the attacks
that we described earlier are out of the scope of this paper.
Also, countermeasures that prevent {\sc sql}
injection attacks~\cite{HVO06}, have already been surveyed.
However, this survey is quite old and since then
the number of countermeasures that detect {\sc sql}
injection attacks alone, has been drastically increased
as we mentioned earlier.

@@

\begin{enumerate}
\item We provide an exploitation model that puts application attacks
  based on source code injection under a common
  perspective.
\item We categorize and analyze proposed defenses using a set of
  criteria that are important for building protection mechanisms.
\item We discuss the issues that arise from our analysis. We put
  emphasis on factors that may hinder widespread adoption of
  approaches and the translation of tools from research to practice.
\end{enumerate}

The rest of the paper is organized as follows.
Section~\ref{sec:attacks} provides the exploitation model
of application attacks based on source code injection.
Section~\ref{sec:dimensions} introduces the dimensions
on which we base the analysis of the various defenses.
Our categorization and analysis is presented in
Section~\ref{sec:defs} and a discussion is provided in
Section~\ref{sec:discussion}. Finally,
Section~\ref{sec:lessons-learned} highlights
some lessons learned from our observations
and Section~\ref{sec:conclusion} concludes the paper.

>>>>

Web application attacks may involve security misconfigurations, broken
authentication and session management, or other issues. Some of the
most dangerous and prevalent web application attacks, however,
exploit vulnerabilities associated with improper validation or filtering
of untrusted inputs, resulting in the injection of malicious
script or domain-specific language code.
Attacks of this type include Cross-Site Scripting ({\sc xss})
attacks~\cite{SG07}, {\sc sql} injection~\cite{RL12b}, and
Cross-Site Request Forgery ({\sc csrf}) attacks~\cite{LZRL09}, among
others.

For the past several years,
these attacks have been topping the lists of the most dangerous vulnerabilities
published by {\sc owasp}~\cite{OWASPtop10},
{\sc mitre}~\cite{MITREtop25}, and other organizations.
For instance, consider the case of {\sc owasp}'s popular Top Ten
project~\cite{OWASPtop10},
which aims to raise awareness about web application security by
identifying some of the most critical risks organizations may face.
In its three consecutive Top Ten lists (2007, 2010, 2013), different source
code-driven injection attacks dominate the top five positions.
At the same time,
attackers find new ways~\cite{HNSHS12,DKH14} to bypass defense
mechanisms using a variety of techniques, despite the numerous
countermeasures that are being introduced. As an example, already by 2006,
there were more than 20 proposed defenses
against {\sc sql} injection attacks~\cite{HVO06}.
Since then, the number has doubled, while researchers have indicated that
the number of {\sc sql} injection attacks has been steadily
increasing in recent years~\cite{SSL12}.

In this paper, we explore how different attacks associated with 
the exploitation of untrusted input validation errors can be modeled 
under a {\it common perspective}. To that end, we propose an exploitation
model which highlights that most of the steps needed to mount
different types of source code injection attacks are
common.\footnote{Note that we do not consider lower-level attacks based on
the exploitation of memory corruption vulnerabilities and the injection
of binary code---see Section~\ref{sec:attacks}.} This is validated by the fact that
there are protection mechanisms that can defend against more than one of these
types of attacks. Then, we categorize a selection of these protection
mechanisms. In our selection we include approaches that
have been proposed in publications cited more than
20 times. We also include some recent mechanisms that have been
presented in top security conferences, even though they have been
cited less than twenty times, so that recent research is not penalized.
Finally, we analyze each
mechanism across the following dimensions:
\begin{itemize}
\item {\it Accuracy:} protection mechanisms are as good
  as their detection capability; this requires low false positive and
  false negative rates.
\item {\it Availability:} whether the mechanism and its
  testbed are publicly available.
\item {\it Computational performance:} the computational
  overhead of the mechanisms at their points of deployment.
\item {\it Ease of use:} whether the protection
  mechanism is practical in terms of deployment
  and can be easily adopted by security experts.
\item {\it Security:} the robustness of the protection mechanism against
  attackers with knowledge of its internals who attempt to circumvent it.
\item {\it Detection Point:} the location where a mechanism detects an attack
  based on our exploitation model.
\end{itemize}

\noindent
All the above requirements are considered important
when building security mechanisms for the protection of
applications~\cite{A01,A00,SPWS13,nature2014}.
Based on this analysis, we identify the advantages
and disadvantages of the
various mechanisms and enumerate some of their common
characteristics. We also draw useful conclusions
about the various protection categories and see how they compare
to each other.

There is already a survey on mitigating software vulnerabilities
in general~\cite{SZ12}. The scope of that
research is much broader and leaves out the majority of the
approaches and mechanisms that we consider in this work.
This is mainly because that work includes countermeasures
that identify vulnerabilities using program analysis
(which takes place during the development or testing phases), while
we consider protection mechanisms that counter
web attacks when they take place. Similarly, static or
dynamic analysis techniques that examine applications to
identify vulnerabilities that may lead to the attacks
that we described earlier are out of scope as well.
While countermeasures against {\sc sql}
injection attacks have been previously surveyed in a 2006 paper~\cite{HVO06},
a significant number of new {\sc sql} injection defenses has been proposed since
then, as we mentioned earlier.

@@

\begin{enumerate}
\item We provide a unified exploitation model for different types of
  web application attacks based on source code injection.
\item We categorize and analyze proposed defenses using a set of
  criteria that are important for building protection mechanisms.
\item We provide insight based on the issues that arise from our analysis. We put
  emphasis on factors that may hinder the widespread deployment of
  protection mechanisms, and the transition of tools from research to practice.
\end{enumerate}

The rest of the paper is organized as follows.
Section~\ref{sec:attacks} presents our proposed model
of web application attacks based on source code injection.
Section~\ref{sec:dimensions} introduces the dimensions
across which we analyze the various defenses.
Our categorization and analysis is presented in
Section~\ref{sec:defs} and a discussion is provided in
Section~\ref{sec:discussion}. Finally,
Section~\ref{sec:lessons-learned} highlights
some lessons learned from our observations
and Section~\ref{sec:conclusion} concludes the paper.
